---
title: "data import"
author: "Steen Flammild Harsted"
date: today
output:
  distill::distill_article:
    self_contained: false
format:
  html: 
    code-fold: true
    code-summary: "Show me the code"
    toc: true
    toc-depth: 2
    number-sections: true
    number-depth: 3
editor: 
  mode: source
execute: 
  eval: false
  echo: true
  message: false
  warning: false
---

```{r setup}
#| include: false
library(tidyverse)
library(here)
library(vroom)
library(naniar)
library(gt)
library(gtsummary)
library(gtExtras)
library(glue)
library(here)
source(here("scripts", "01_import.R"))
```

# Data import

<br><br>

#### Download the `data_day_3.zip` file from ItsLearning and unzip it in your `raw_data` directory
The unzipped folder contains:

*  A .csv file called "Kronisk smerte - udvikling.csv". It has been exported by a Danish version of Excel. Is this a good name for a file?
*  A set of files related to a motioncapture project
  -  A Stata file (.dta) containing the sex of the participants
  -  A .csv file containing the age (in months) of the partipants. 
  -  19 simplified motioncapture files of children performing vertical jumps
     - `Frame` and `Time_Seconds` are time variables
     - `CGY` gives you the height (in mm) of their center of gravity
*  A folder called `challenge` that you need if you want to solve the challenge assignment

<br><br>

#### Import "kronisk smerte - udvikling.csv" 
* `read.csv()`
* `read.csv2()`
* `read_csv2()`
*  What is the difference in the output? Why?
```{r}
#| output: false
path <- here("raw_data", "data_day_3"  ,"Kronisk smerte - udvikling.csv")
read.csv(path)
read.csv2(path)
read_csv2(path)
```

<br><br>

#### Import the files `id_age.csv` and `id_sex.dta`, combine them (use `full_join()`), and assign the combined dataframe to an object
*  The `tidyverse` includes the `haven` package that can read Stata´s .dta files
  - The function to use for the .dta file is `haven::read_dta()`
*  Investigate the two files before you combine them. Do you need to change anything?
*  The `dplyr` function `full_join()` will help you to combine the two imported objects.

```{r}
#| output: false
a <- haven::read_dta(here("raw_data", "data_day_3", "id_sex.dta"))

b <- read_csv2(here("raw_data", "data_day_3","id_age.csv")) %>%
     rename(ID = id)  # Rename the id column

df_descriptives <- full_join(a, b)
rm(a,b) # Remove the objects a and b so that your environment is less crowded
```

<br><br>

#### Use `list.files()` to generate an object called `files` containing the filenames of the 19 motioncapture files

<details><summary>Hints for the `pattern = `argument in `list.files()`</summary>
```{r}
#| eval: false
#| code-fold: false
, pattern = NULL          # The default setting. List all files in our directory
, pattern = ".csv$"       # all files in our directory that ends with ".csv"
, pattern = "^desc"       # all files in our directory that starts with "desc"
, pattern = "[0-9].csv$"  # all files in our directory that ends with "[a number from 0-9].csv"
```
</details>

```{r}
#| output: false
files <- list.files(here("raw_data", "data_day_3"), pattern = "[0-9].csv$")
```

<br><br>

#### Import the 19 motioncapture files 

* Use `read_csv()``
* The `file = ` should include the path to the folder (use `here()`), and the `files` object you just created.
```{r}
#| output: false
df_mocap <- read_csv(here("raw_data", "data_day_3", files))
```

*  Combine the motion capture files with the descriptives object you created before
```{r}
#| output: false
df_all <- full_join(df_descriptives, df_mocap)
```

<br><br>

#### Update the object you just created 
*  Change ID to a factor
*  Change sex to a factor with `levels = c(1,2), labels = c("Boy", "Girl")`
*  Change age from months to years
*  Retain the row with the highest value of `CGY` for each of the children
```{r}
#| output: false
df_all <- df_all %>% 
  mutate(
    ID = factor(ID),
    sex = factor(sex, levels = c(1,2), labels = c("Boy", "Girl")),
    age = age/12) %>% 
  group_by(ID) %>% 
  filter(CGY == max(CGY, na.rm = TRUE))
```

<br><br>

#### Save the object you just created
*  Save the object you created before as an ".RData" file, save it in your "clean_data" folder
*  Save the object you created before as an ".csv" file, save it in your "clean_data" folder
*  What are the pros and cons of the two file types?
```{r}
#| output: false
save(df_all, file = here("clean_data", "my_data.RData"))
write_csv(df_all, file = here("clean_data", "my_data.csv"))
```

<br><br>

### LOOKING FOR A CHALLENGE?

#### Inspect the 19 motioncapture files in the folder called `challenge` 
*  What important piece of information is missing from the data in the files?
*  Where can you find this data? <details><summary>Answer</summary> The ID can only be found in the filename. You need to find a way to piece the ID together with the data in files </details>
  
#### Install the package `vroom` and load it
*  `vroom` is a very fast package for importing `.csv` files. (hence the name)  
*  The main function in the `vroom` package is `vroom()`
*  `vroom()` has the argument `delim = ` that allows you specify the delimter you want
*  Read the documentation for `vroom()`
      - What does the `id` argument in the `vroom()` function do?

<br><br>  

####  Use `vroom()` to import all the mocap files in the challenge folder
  - Use the `id = ` argument in the `vroom()` function.

```{r}
#| output: false
files_chal <- list.files(here("raw_data", "data_day_3", "challenge"), pattern = "[0-9].csv$")
df_mocap_chal <- vroom::vroom(here("raw_data", "data_day_3", "challenge", files_chal),
                              id = "filename")
```

<br><br>

#### We need to extract the ID from the filename column now
*  Create a column called `ID` that only contains the `ID` part from the filename
*  You need to use regular expressions to solve this
*  Two functions that may help you are `str_extract()` and `str_remove()`
  - Both functions have a `pattern = ` argument that must be a regular expression
  - Use them inside a `mutate()` call
```{r}
#| output: false

df_mocap_chal <- df_mocap_chal %>%
  mutate(ID = str_extract(filename, pattern = "[0-9]+_3.1"),  # Capture digits that come before _3.1
         ID = str_remove(ID, pattern = "_.+$"),               # Remove the _3.1 part
         ID = as.numeric(ID))                                 # Change ID from a string to a numeric variable. You can also change it to a factor
```


