---
title: "quarto documents"
author: "Steen Flammild Harsted"
date: '`r Sys.Date()`'
output:
  distill::distill_article:
    self_contained: false
creative_commons: CC BY
format:
  html: 
    code-fold: true
    code-summary: "Show me the code"
    toc: true
    toc-depth: 2
    number-sections: true
    number-depth: 2
editor: 
  mode: source
execute: 
  output: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


\ \ \ 

# quarto documents

<br><br>

### Start a new .qmd document
 File ->  
 New File ->  
 Quarto Document ->  
 * Make sure the settings are set to "Document" and "HTML" (both are default settings), 
 * Uncheck 'Use visual markdown editor'
 * press "Create" 
 
 The document contains both R and Markdown code.   
 Save your file  (Important: file_name_matters)
 Pres "Render" and see what happens.

    
 <br><br>

### Markdown Code
[Markdown code basics](https://quarto.org/docs/authoring/markdown-basics.html)
    
### Headlines
Headlines are made using one or more # signs followed by a space  
\# Largest headline  
\## Second largest headline  
\### Third largest headline  
\##### Smallest headline possible.   

* Practice making different headlines.  

<br><br>

### **Bold** and *Italic*
\*text\*  becomes *text*  
\*\*text\*\*  becomes **text**  
\*\*\*text\*\*\*  becomes ***text***  

* Practice writing something in *italics* and **bold**  

<br><br>
  
### New lines
To break a line in R Markdown and have it appear in your output, use two trailing spaces and then hit return.  
If you want a hard line break end your lines with a `\`

* Practice making line-breaks.
  
<br><br>
  
### Code chunk
Insert a code chunk in your document (Default `CTRL+Alt+I`)  
If pressing CTRL+Alt+I doesn't work click Tools -> Command Palette -> and type "Chunk" in the search field  
  
What you write inside the codechunk will be evaluated as R code.

Write 2+2 inside the Code chunk you just created, and Knit the document.
  
2+2 equals
```{r}
2+2
```
  
<br><br>
  
### Run Code chunk
You don't need to Knit your document in order to see the output of your code.  
Place the cursor inside the code chunk you just made and Click the green triangle (play-sign) in the right side of the code chunk.  
Notice the output in the console.  
  
R comes with many inbuilt datasets, one of them is the `iris` dataset.  
Plot the `iris` dataset.  
Insert a codechunk, and write `plot(iris)` inside the codechunk.  
Click the green triangle.  
 
**Lifehack:**   
pressing `ALT+Enter` will run the current line or selection.  
pressing `CTRL+Enter` will run the current line or selection and jump to next line.  
pressing `CTRL+ALT+C` will run the current code chunk (instead of just the current line).  


  
<script src="https://giscus.app/client.js"
        data-repo="sorenoneill/r4phd"
        data-repo-id="R_kgDOJ9etDQ"
        data-category="Announcements"
        data-category-id="DIC_kwDOJ9etDc4CYApF"
        data-mapping="pathname"
        data-strict="0"
        data-reactions-enabled="0"
        data-emit-metadata="0"
        data-input-position="bottom"
        data-theme="cobalt"
        data-lang="en"
        crossorigin="anonymous"
        async>
</script>