---
title: "Missings"
subtitle: "and the `forcats` package"
author: "Steen Flammild Harsted"
date: today
output:
  distill::distill_article:
    self_contained: false
format:
  html: 
    code-fold: true
    code-summary: "Show me the code"
    toc: true
    toc-depth: 2
    number-sections: true
    number-depth: 3
editor: 
  mode: source
execute: 
  eval: false
  echo: true
  message: false
  warning: false
---

```{r setup}
#| include: false
library(tidyverse)
library(here)
library(vroom)
library(naniar)
library(gt)
library(gtsummary)
library(gtExtras)
library(glue)
library(here)
source(here("scripts", "01_import.R"))
```

## Missing Data

<br><br>

#### Install the `naniar` package and load it
Use `install.packages()` to download the `naniar` package. When you have done that add `library(naniar)` to the code chunk where you call your libraries. Execute the line. 


<br><br>

#### Explore missing values in `starwars` using
*  `vis_miss()` 
*  `gg_miss_upset()`
*  What happens if you add `, nsets = 6` inside `gg_miss_upset()`
*  `naniar` contains more useful functions. You can see them [here](https://cran.r-project.org/web/packages/naniar/vignettes/naniar-visualisation.html)
```{r}
#| output: false
vis_miss(starwars)
gg_miss_upset(starwars)
```
