---
title: "Extra Material"
author: "Steen Flammild Harsted"
date: today
output:
  distill::distill_article:
    self_contained: false
format:
  html: 
    code-fold: true
    code-summary: "Show me the code"
    toc: true
    toc-depth: 2
    number-sections: true
    number-depth: 2
editor: 
  mode: source
execute: 
  eval: false
  echo: true
  message: false
  warning: false
---
```{r}
#| include: false
library(tidyverse)
library(here)
```



## Diving Deeper into ggplot2 Extensions {.unnumbered}
`ggplot2 ` stands as a cornerstone for data visualization in R. But the richness of the R ecosystem extends far beyond its foundational tools. Complementing the core offerings of `ggplot2` is a suite of powerful extension packages, each designed to elevate and refine your visualization experiences.

In this section, we'll journey through three particularly impactful extensions: `gganimate`, `ggiraph`, and `ggside`.


::: {.callout-tip}
## Twist!
While our exploration in this section will be guided, it comes with a twist. We'll be immersing ourselves in blogs and webpages from the broader R community. The aim? To prepare you for a future where you will explore the R ecosystem independently.
:::

Here is a brief introduction to three packages. Start with the one you find most interesting.

### `gganimate`
Breathe life into your static plots by transforming them into animations. Whether you're showcasing trends over time or simply adding a dynamic touch to your visuals, gganimate is your go-to tool.

```{r}
#| eval: true
#| echo: false
library(gganimate)
library(gapminder)
ggplot(gapminder::gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  facet_wrap(~continent) +
  # Here comes the gganimate specific bits
  labs(title = 'Year: {frame_time}', x = 'GDP per capita', y = 'life expectancy') +
  transition_time(year) +
  ease_aes('linear')
```


### `ggiraph` 
Make your plots interactive! With `ggiraph`, you can create hover-over effects, clickable elements, and other interactive features that make your plots more engaging and informative.

```{r}
#| eval: true
#| echo: false
#| output: true

library(ggiraph)
data <- mtcars
data$carname <- row.names(data)

gg_point = ggplot(data = data) +
    geom_point_interactive(aes(x = wt, y = qsec, color = disp,
    tooltip = paste("Model:", carname, "\nHorsepower:", hp, "\nCylinders:", cyl), 
    data_id = carname)) + 
  theme_minimal()+
  labs(
    title = "Try to hover your mouse over the points",
    x = "Weight (1000 lbs)",
    y = "1/4 mile time")

girafe(ggobj = gg_point)
```



### `ggside`
Expand the canvas of your data stories by crafting marginal plots alongside your primary visualization. With `ggside`, you can seamlessly incorporate histograms, density plots, or boxplots directly adjacent to your main plot. Whether it's enriching a scatterplot with marginal distributions or adding supplementary insights, `ggside` offers that extra dimension to your visual narratives.

```{r}
#| eval: true
#| echo: false
library(ggside)
ggplot(mpg, aes(displ, hwy, colour = class)) + 
  geom_point(size = 2) +
  geom_xsideboxplot(aes(y =class), orientation = "y") +
  geom_ysidedensity(aes(x = after_stat(density)), position = "stack") +
  theme(ggside.panel.scale = .3) +
  scale_xsidey_discrete() +
  scale_ysidex_continuous(guide = guide_axis(angle = 90), minor_breaks = NULL) 
```

<br><br><br>

## `gganimate`

Most packages have a webpage where the package and its core functions are introduced. Head the `gganimate` website and read the [getting started](https://gganimate.com/articles/gganimate.html#your-first-animation) section. Then continue to exercises here where we will practice using the `soldiers` dataset that you already know.

You need to install `gganimate`

::: {.callout-tip collapse=true}
#### Installing `gganimate`
```{r}
#| eval: false
#| code-fold: false
install.packages("gganimate")
```
:::

```{r}
#| eval: true

# Load libraries
library(tidyverse)  # Data wrangling and plots
library(here)       # File control in project
library(gganimate)  # Animate plots

# Load the cleaned soldiers dataset
source(here("scripts", "01_import.R"))
```

<br>

### Animate this plot so its shows difference between men and women
```{r}
#| code-fold: false
#| output: true
#| eval: true
soldiers %>% 
  ggplot(aes(x = weightkg, y = Installation, fill = Installation))+
  geom_boxplot(outlier.shape = NA,
               alpha = 0.5)
```

*  Assign the animation to an object 

```{r}
# One option
my_animation <- soldiers %>% 
  ggplot(aes(x = weightkg, y = Installation, fill = Installation))+
  geom_boxplot(outlier.shape = NA,
               alpha = 0.5)+
  transition_states(sex,
                    transition_length = 1,
                    state_length = 2)+
  labs(title = "Weight in Kgs of male and female soldiers",
       subtitle = "Now showing {closest_state}")

my_animation
```

<br>

```{r}
# Another option
soldiers %>% 
  ggplot(aes(x = weightkg, y = Installation, fill = Installation, seq_along = sex))+
  geom_boxplot(outlier.shape = NA,
               alpha = 0.5)+
  labs(title = "Weight in Kgs of male and female soldiers",
       subtitle = "Now showing {closest_state}")+
  transition_states(sex,
                    transition_length = 0.5,
                    state_length = 2)+
  enter_fade()+
  exit_fly()
```

<br>

### Save your animation in the plots folder you created earlier

```{r}
#| eval: false
anim_save(
    filename = here("plots", "my_animation.gif"),
    animation = my_animation
  )
```

<br>

### `Oxboys` data

Load the`nlme` package and examine the `Oxboys` dataset that it comes with.  
You might need to install the `nlme` package first.
```{r}
library(nlme)
head(Oxboys)
?Oxboys
```

### Create an animation that shows how `height` changes with `age`
*  Use `transition_reveal()`
```{r}
Oxboys %>% 
  ggplot(aes(x = age, y = height))+
  # geom_boxplot(fill = "steelblue", 
  #              alpha = 0.5, 
  #              width = 0.3,
  #              outlier.shape = NA)+
  geom_line(aes(group = Subject), color = "grey", alpha = 1)+
  geom_point(aes(group = Subject))+
  labs(x = "Standardized age \n(years)",
       y = "Height\n(cm)")+
  theme_classic()+
  transition_reveal(age)
```


<br><br><br>

## `ggiraph`
Another useful resource for learning R is blogposts. These are often written in a much more accesible language. For learning `ggiraph` we will use a blogpost written by [Albert Rapp](https://albert-rapp.de/). Head to his [blogpost](https://albert-rapp.de/posts/ggplot2-tips/17_ggiraph/17_ggiraph.html) about `ggiraph` and read it. You don't need to read the sections about `shiny`apps. When you have done that continue to exercises below.


You need to install `ggiraph`. 

::: {.callout-tip collapse=true}
#### Installing `ggiraph`
```{r}
#| eval: false
#| code-fold: false
install.packages("ggiraph")
```
:::

<br>

### Load the `tidyverse`, `here`, `ggiraph`, `patchwork`, and the cleaned `soldiers` dataset
```{r}
#| eval: true

# Load libraries
library(tidyverse)  # Data wrangling and plots
library(here)       # File control in project
library(ggiraph)    # Interactive plots
library(patchwork)  # Combining plots

# Load the cleaned soldiers dataset
source(here("scripts", "01_import.R"))
```

<br>

### Explore `Fort Hood`, `Camp Arnold`, and `Fort Doughnut`
*  Filter you data to the three camps
*  Make a interactive boxplot that shows the name of the installation

::: {.callout-tip collapse=true}
#### Hint
You can use the base R functions `paste()` or `paste0()` to combine text and values in your tooltip.  
You can also use `glue()` from the `glue` package, but you need to install that package first. In the blogpost Albert uses the `glue()` function.
:::

::: {.callout-tip collapse=true}
#### Size of your interactive plot
With the default options, the interactive plots tends to be oversized. Add this option to reduce it
```{r}
#| eval: false
#| code-fold: false
girafe(ggobj = my_plot,
       options = list(opts_sizing(rescale = FALSE)))
```

:::


```{r}
my_plot <- soldiers %>% 
  filter(Installation %in% c("Fort Hood", "Camp Arnold", "Fort Doughnut")) %>% 
  ggplot(aes(x = BMI, y = Installation, fill = Installation))+
  geom_boxplot_interactive(
    aes(tooltip = paste0("This is the ", Installation, " installation")))+
  labs(y = NULL)+
  theme_classic()

girafe(ggobj = my_plot)
```

### Add an interactive geom_jitter layer.
* facet the plot on sex
* Make a tooltip that gives you the subjectid, BMI, height, weight, biceps circumference, and waist circumference.

::: {.callout-tip collapse=true}
#### Hint line breaks
"\\n" Gives you a line break
:::

```{r}
my_plot <- soldiers %>% 
  filter(Installation %in% c("Fort Hood", "Camp Arnold", "Fort Doughnut")) %>% 
  ggplot(aes(x = BMI, y = Installation, fill = Installation))+
  facet_wrap(~sex, ncol = 1)+
  labs(y = NULL)+
  geom_boxplot_interactive(
    aes(tooltip = paste0("This is the ", Installation, " installation")),
    outlier.shape = NA) +
  geom_jitter_interactive(
    aes(tooltip = paste0("Subjectid: ", subjectid,
                         "\n BMI:   ", round(BMI), " (", category, ")",
                         "\nHeight: ", round(heightcm), " cm",
                         "\nWeight: ", round(weightkg), " kg",
                         "\nBiceps: ", round(bicepscircumference), " mm", 
                         "\nWaist:  ", round(waistcircumference), " mm")),
    height = 0.2,
    alpha = 0.3,
    shape = 21
  )+
  theme_classic()

girafe(ggobj = my_plot,
       options = list(opts_sizing(rescale = FALSE)))
```

### Create another interactive plot with `geom_jitter_interactive()`
*  Combine your new plot with the one you made above using `patchwork`
*  Set the data_id argument to: `data_id = subjectid` for the points
*  Use the options that Alfred suggested in his post (see tip below)

::: {.callout-tip collapse=true}
#### Options
```{r}
#| code-fold: false
girafe(ggobj = my_plot,
       options = list(
         opts_hover(css = ''),
         opts_hover_inv(css = "opacity:0.2;"), 
         opts_sizing(rescale = FALSE)
  ))
```
:::


```{r}
my_plot_a <- soldiers %>% 
  filter(Installation %in% c("Fort Hood", "Camp Arnold", "Fort Doughnut")) %>% 
  ggplot(aes(x = BMI, y = Installation, fill = Installation, ))+
  facet_wrap(~sex, ncol = 1)+
  labs(y = NULL,
       x = NULL,
       title = "BMI")+
  geom_boxplot_interactive(
    aes(data_id = Installation,
        tooltip = paste0("This is the ", Installation, " installation")),
    outlier.shape = NA) +
  geom_jitter_interactive(
    aes(data_id = subjectid,
        tooltip = paste0("Subjectid: ", subjectid,
                         "\n BMI:   ", round(BMI), " (", category, ")",
                         "\nHeight: ", round(heightcm), " cm",
                         "\nWeight: ", round(weightkg), " kg",
                         "\nBiceps: ", round(bicepscircumference), " mm", 
                         "\nWaist:  ", round(waistcircumference), " mm")),
    height = 0.2,
    alpha = 0.5,
    shape = 21
  )+
  theme_classic()

my_plot_b <- soldiers %>% 
  filter(Installation %in% c("Fort Hood", "Camp Arnold", "Fort Doughnut")) %>% 
  ggplot(aes(x = waistcircumference, y = Installation, fill = Installation))+
  facet_wrap(~sex, ncol = 1)+
  labs(y = NULL,
       x = NULL,
       title = "Waist circumference")+
  geom_boxplot_interactive(
    aes(data_id = Installation,
        tooltip = paste0("This is the ", Installation, " installation")),
    outlier.shape = NA) +
  geom_jitter_interactive(
    aes(data_id = subjectid,
        tooltip = paste0("Subjectid: ", subjectid,
                         "\n BMI:   ", round(BMI), " (", category, ")",
                         "\nHeight: ", round(heightcm), " cm",
                         "\nWeight: ", round(weightkg), " kg",
                         "\nBiceps: ", round(bicepscircumference), " mm", 
                         "\nWaist:  ", round(waistcircumference), " mm")),
    height = 0.2,
    alpha = 0.5,
    shape = 21
  )+
  theme_classic()

my_plot <- my_plot_a + my_plot_b +
  plot_layout(guides = "collect") &
  theme(legend.position='bottom')

girafe(ggobj = my_plot,
       options = list(
         opts_hover(css = ''),
         opts_hover_inv(css = "opacity:0.2;"), 
         opts_sizing(rescale = FALSE)
  ))
```




<script src="https://giscus.app/client.js"
        data-repo="sorenoneill/r4phd"
        data-repo-id="R_kgDOJ9etDQ"
        data-category="Announcements"
        data-category-id="DIC_kwDOJ9etDc4CYApF"
        data-mapping="pathname"
        data-strict="0"
        data-reactions-enabled="0"
        data-emit-metadata="0"
        data-input-position="bottom"
        data-theme="cobalt"
        data-lang="en"
        crossorigin="anonymous"
        async>
</script>