---
title: "Comments - where to"
author:
  - name: Søren O'Neill
    url: www.oneill.dk
date: '`r Sys.Date()`'
output:
  distill::distill_article:
    self_contained: false
#collections:
#  posts:
#    disqus: reproducible-finance-with-r
#categories:
#  - Installation
creative_commons: CC BY
format: html
editor: source
---

```{r include=FALSE}
#library(vembedr)
library(tidyverse)
library(ggplot2)
library(DiagrammeR)
```

# Where to add comments in R code and Quarto MarkDown

::: {.callout-note collapse=false}
# Exercise #1

What do you think about the way these comments have been placed? Will it work?

```{r}
#| eval: false
#| echo: true

# this comment definitely okay
x <- 2 + 2

# ...but is the following okay?
tibble(x=1:10, y=2:11) %>% # ..is it okay, mid-pipe?
  filter(y>5 & # and how about breaking in the middle..
  x != 9) # ..of a statement like 'filter'?
```

:::

::: {.callout-note collapse=true}
# Exercise #1 Solution

This is what the output looks like:

```{r, echo=FALSE}
#|  eval: true
#|  echo: false

# this comment definitely okay
x <- 2 + 2

# ...but is the following okay?
tibble(x=1:10, y=2:11) %>% # ..is it okay, mid-pipe?
  filter(y>5 & # and how about breaking in the middle..
  x != 9) # ..of a statement like 'filter'?
```

...so the code seems to run without fault - R is actually quite tolerant about placing comments here, there and everywhere, but is it good practice, though?

These comments are syntactically okay ... but did they make your code *simpler* for humans to read? No not really. In fact, they made it rather more difficult than it needed to be. So click below to keep it simple:

:::

::: {.callout-note collapse=true}
## Keep it simple!

```{r}
#| eval: false
#| echo: true
# Comment lines before the code
# can be multiple lines and does
# not disturb your reading of the
# code ... best practice!

x <- 2 + 2 # use inline comments sparringly
y <- x^2 - c(2, 5, 8, 10:14) # ..they are distracting!
```

:::


<script src="https://giscus.app/client.js"
        data-repo="sorenoneill/r4phd"
        data-repo-id="R_kgDOJ9etDQ"
        data-category="Announcements"
        data-category-id="DIC_kwDOJ9etDc4CYApF"
        data-mapping="pathname"
        data-strict="0"
        data-reactions-enabled="0"
        data-emit-metadata="0"
        data-input-position="bottom"
        data-theme="cobalt"
        data-lang="en"
        crossorigin="anonymous"
        async>
</script>