---
title: "Load data from a CSV file"
description: |
  How to load data into R from a _comma Separated Values_ file.
author:
  - name: Søren O'Neill
    url: www.oneill.dk
date: "2023-11-26"
#output:
#  distill::distill_article:
#    self_contained: false
#collections:
#  posts:
#    disqus: reproducible-finance-with-r
categories:
  - RStudio
  - Input
creative_commons: CC BY
format: html
editor: source
number-sections: true
---

## What is a CSV file?

* CSV is an acronym for '_Comma Separated Values_'.
* It is a format for storing data in a simple text file.
* CSV files inherit all the benefits of working with simple text files.
* Conversely, CSV files are limited in the complexity of the data it can store: only simple tabular data.
* Read more about CSV files here [wiki](https://en.wikipedia.org/wiki/Comma-separated_values). 

## An example file

An example of a CSV file could look like this:

```{r}
#| eval: false
#| echo: true
#| code-line-numbers: false
id,test,value,group,note
1,1,12,"A",""
1,2,14,"B",""
2,1,15,"C","My note"
2,2,13,"A",""
```

::: {.callout-tip collapse="true" title="Think! What does the CSV file tell us about the data? ...and not?"}
From the raw text of the CSV file we can deduce, that the data consists of 5 variables and 4 observations, delimited by '**,**'. We can take a guess as to the meaning of the variables from their names i line 1, but we can not know. Also, **note** that a CSV file does not necessarily include a title line.

Also, ask yourself 

* _What is the data type of each variable?_ (integer, numerical, factors/categorical, etc)? 
* _What are the allowed or valid values for each variable?_
* _What happens if a variable is missing_

...such meta-data is not stored within a CSV and we need to keep tabs on it ourselves: It is a good idea to maintain a _data definition_ file along with the CSV file itself, which described the data in more detail.
:::

### Is a comma the best choice?

In fact, as decimal numbers (like $\frac14$) are typed as _0.25_ in some locales/countries and _0,25_ in others, there is a risk that decimal commas can be mistaken for value-delimiting commas, and vice versa. It is thus not uncommon to use a semi-colon instead of a comma.

## Tidyverse solution

* Tidyverse includes the package [readr](https://readr.tidyverse.org/) which provides a number of functions to read CSV files. 
* The function `read_delim()` is the generic function which reads tabular data delimited by whatever character is specified in the function call.
* The functions `read_csv()`, `read_tsv()` and `read_csv2()` are just special cases of `read_delim()` (using comma, tab and semi-colon as delimiters).

::: {.callout-tip collapse="true" title="If you need a base-R alternative.."}
Base R includes the function `read.table()` which has a number of special cases, including `read.csv()` that also serves to read tabular data from CSV files. **Note** the use of dot rather than an underscore in the function name.
:::

## RTFM -- Read the fine manual

Firstly, acquaint yourself with the `read_delim()` function using the [help function](https://www.r-project.org/help.html) in RStudio. 

## Try it out

Ideally, use you own CSV data file for these exercises.

Alternatively, you will find a [sample CSV file](./example_csv_file.csv) here

* Inspect the data with `my_dataframe` and `str(my_dataframe)`.

::: {.callout-tip collapse="true" title="Coding example.."}

```{r}
#| eval: false
#| echo: true

my_data <- read_csv(here("path_to_file", "fil.csv"))
my_data
str(my_data)
```
:::


## Exercises

* What happens if you use `read_csv()` and `read_csv2()` respectively without specifying a delimiter?
* What happens if you set the parameter `col_names = FALSE`?
* What happens if you set the parameter `na = "NA"`?
* Look at the data structure as revealed by the `str()` function, are all you variables of the type you expected them to be?
    - How could you use the parameter `col_types` to specify data types?

<script src="https://giscus.app/client.js"
        data-repo="sorenoneill/r4phd"
        data-repo-id="R_kgDOJ9etDQ"
        data-category="Announcements"
        data-category-id="DIC_kwDOJ9etDc4CYApF"
        data-mapping="pathname"
        data-strict="0"
        data-reactions-enabled="0"
        data-emit-metadata="0"
        data-input-position="bottom"
        data-theme="cobalt"
        data-lang="en"
        crossorigin="anonymous"
        async>
</script>