---
title: "Base R - missing data"
author: "Søren O'Neill"
date: today
output:
  distill::distill_article:
    self_contained: false
creative_commons: CC BY
format: html
editor: 
  mode: source
---

# Handling missing data

Take a look at this simple code:

```{r}
# Our seven participants self-reported their age:
participants_age <- c(23,25,23,21,25,NA,20)

# What is the mean age?
mean(participants_age)
```

The **NA** values indicates a data point that is _Not Available_, i.e. missing. This is fundementally _not_ the same as NULL or 0. So how would you calculate a mean value? Well, when the age of one participant is unavailable, so is the mean value and R, quite rightly, reports the mean as **NA**.

You might think, it would be obvious to calculate the mean value of the available values and ignore the one NA. In some cases, that makes sense and is a valid decision to make (or perhaps not). The important think is: _you have to decide how to deal with missing values - R won't do it for you_. This is not a bug, it's a (safety) feature.

Look at this amended code:

```{r}
# Our seven participants self-reported their age:
participants_age <- c(23,25,23,21,25,NA,20)

# What is the mean age?
mean(participants_age, na.rm=TRUE)
```

What is your best quess as to the meaning of `na.rm=TRUE`?

<script src="https://giscus.app/client.js"
        data-repo="sorenoneill/r4phd"
        data-repo-id="R_kgDOJ9etDQ"
        data-category="Announcements"
        data-category-id="DIC_kwDOJ9etDc4CYApF"
        data-mapping="pathname"
        data-strict="0"
        data-reactions-enabled="0"
        data-emit-metadata="0"
        data-input-position="bottom"
        data-theme="cobalt"
        data-lang="en"
        crossorigin="anonymous"
        async>
</script>