---
title: "Reproducible research with R -- workflow #1"
author:
  - name: Søren O'Neill
    url: www.oneill.dk
date: '`r Sys.Date()`'
output:
  distill::distill_article:
    self_contained: false
#collections:
#  posts:
#    disqus: reproducible-finance-with-r
#categories:
#  - Installation
creative_commons: CC BY
format: html
editor: source
---

```{r include=FALSE}
#library(vembedr)
library(tidyverse)
library(ggplot2)
library(DiagrammeR)
```


# Reproducible Research

Let's say you have noticed a wrong entry in your data set ... in invalid CPR number. You might decide to delete the observation from your data like this:

```{r}
#| eval: false
#| echo: true

data <- read.csv("my_data_file.csv")
# excluded as participant entered an invalid CPR number
data <- data %>% filter(id != "2321369-1212")
```

_...but why not just delete that observation from the csv file?_

:::{.callout-note collapse=true}

## Reproducible Research

The principles of Reproducible Research admonish you that the **entire process** from raw data to output should be:

- scripted
- reproducible
- reversible

What happens when at 6 months later, someone notices that you downloaded 106 rows of data from RedCap, but your manuscript reports n=105? Do you think you'll remember then?

Or for a more dramatic example, look at this youtube video {{< video https://youtu.be/d2Tm3Yx4HWI >}} when you have the time (~20 minutes)

:::

    
<script src="https://giscus.app/client.js"
        data-repo="sorenoneill/r4phd"
        data-repo-id="R_kgDOJ9etDQ"
        data-category="Announcements"
        data-category-id="DIC_kwDOJ9etDc4CYApF"
        data-mapping="pathname"
        data-strict="0"
        data-reactions-enabled="0"
        data-emit-metadata="0"
        data-input-position="bottom"
        data-theme="cobalt"
        data-lang="en"
        crossorigin="anonymous"
        async>
</script>