---
title: "Comments - what to"
author:
  - name: Søren O'Neill
    url: www.oneill.dk
date: '`r Sys.Date()`'
output:
  distill::distill_article:
    self_contained: false
#collections:
#  posts:
#    disqus: reproducible-finance-with-r
#categories:
#  - Installation
creative_commons: CC BY
format: html
editor: source
---

```{r include=FALSE}
#library(vembedr)
library(tidyverse)
library(ggplot2)
library(DiagrammeR)
```

# Meaningful comments

What comments would be relevant here?

::: panel-tabset

# Question

Suggest comments for this code... think 'what' and 'why'

```{r eval=FALSE, echo=TRUE}
# 1
d <- read.csv("my_data_file.csv")

# 2
d <- d %>% filter(id != "241269-1212")

# 3
d <- d %>% 
  mutate(s=factor(c("M", "F"))[as.numeric(substr(id,nchar(id),nchar(id))) %% 2])

```


# Solution

Suggestions for meaningful comments

```{r eval=FALSE, echo=TRUE}
# ?? comments necessary ??
d <- read.csv("my_data_file.csv")

# excluded as participant entered an invalid CPR number
d <- d %>% filter(id != "2321369-1212")

# set 's' to F(emale) or M(ale) depending on odd/even id number in CPR
d <- d %>% 
  mutate(s=factor(c("M", "F"))[as.numeric(substr(id,nchar(id),nchar(id))) %% 2])

```
::::

## Commenting your code

**Main points**

* **Always** comment your code
* Comments should explain your thinking : 'why'
* With good code the 'how' is self-evident
    - ..but if your code can not be self-evident: explain in comments


<script src="https://giscus.app/client.js"
        data-repo="sorenoneill/r4phd"
        data-repo-id="R_kgDOJ9etDQ"
        data-category="Announcements"
        data-category-id="DIC_kwDOJ9etDc4CYApF"
        data-mapping="pathname"
        data-strict="0"
        data-reactions-enabled="0"
        data-emit-metadata="0"
        data-input-position="bottom"
        data-theme="cobalt"
        data-lang="en"
        crossorigin="anonymous"
        async>
</script>